---
title: "logoSTS"
---

> **Superseded.** The production logo and lockup assets are now generated by
> `logoSTS2.qmd` (fitted-bézier SVGs) and `lockup.qmd` (logo + wordmark).
> This file is kept as the original design source — the geometry the logo was
> born from. The gridSVG/spline export chunks it once had were removed
> (July 2026); they are in git history if ever needed.

```{r}
library(tidyverse)
theme_set(theme_void())
```

```{r}
col1 = "#99D9EA"
col2 = "#00A2E8"
```

```{r}
# nx = 1e6  # old: dense polylines (59 MB SVG!)
nx = 500    # enough for accurate crossing detection in the plots below
```

```{r}


dat <- tibble(
  x = seq(-3, 3, length = nx),
  y = dnorm(x),
  y2 = dnorm(x, .6, .7)/2,
  y3 = dnorm(x, -.6, .7)/2,
  )

dat
```

```{r}
dat_poly = rbind(
  dat |> filter(row_number() <= which.max(y2 > y)) |> transmute(x, y = y2),
  dat |> filter(row_number() <= which.max(y2 > y)) |> transmute(x, y) |> arrange(desc(x))
)

dat_poly2 = rbind(
  dat |> filter(row_number() >= last(which(y3 > y2))) |> transmute(x, y = y3),
  dat |> filter(row_number() >= last(which(y2 > y))) |> transmute(x, y = y) |> arrange(desc(x)),
  dat |>
    filter( row_number() >= last(which(y3 > y2)) ) |>
    filter( row_number() <= last(which(y2 > y)) ) |>
    transmute(x, y=y2) |> arrange(desc(x))
)

```

# Full fill no border

```{r}
#| eval: false
#| fig-height: 4
#| fig-width: 6
glogo_poly <- ggplot() +

  geom_polygon(
    data = dat_poly,
    mapping = aes(x=x, y=y),
    fill = col1,
    colour = NA
  ) +

  geom_polygon(
    data = dat_poly2,
    mapping = aes(x=x, y=y),
    fill = col2,
    colour = NA
  ) +

  geom_polygon(
    data = dat_poly,
    mapping = aes(x=-x, y=-y),
    fill = col1,
    colour = NA
  ) +

  geom_polygon(
    data = dat_poly2,
    mapping = aes(x=-x, y=-y),
    fill = col2,
    colour = NA
  )

glogo_poly
```

```{r}
#| eval: false
ggsave("gglogo_poly.pdf", width = 3, height = 2)
ggsave("gglogo_poly.svg", width = 3, height = 2)
ggsave("gglogo_poly.emf", width = 3, height = 2)
```

# White border

```{r}
#| eval: false
#| fig-height: 2
#| fig-width: 3

lw = 1.1
col = "white"

glogo_poly_b <- ggplot() +

  geom_polygon(
    data = dat_poly,
    mapping = aes(x=x, y=y),
    fill = col1,
    colour = col,
    linewidth=lw
  ) +

  geom_polygon(
    data = dat_poly2,
    mapping = aes(x=x, y=y),
    fill = col2,
    colour = col,
    linewidth=lw
  ) +

  geom_polygon(
    data = dat_poly,
    mapping = aes(x=-x, y=-y),
    fill = col1,
    colour = col,
    linewidth=lw
  ) +

  geom_polygon(
    data = dat_poly2,
    mapping = aes(x=-x, y=-y),
    fill = col2,
    colour = col,
    linewidth=lw
  )

glogo_poly_b
```

```{r}
#| eval: false
ggsave("gglogo_poly_w.pdf", width = 3, height = 2)
ggsave("gglogo_poly_w.svg", width = 3, height = 2)
ggsave("gglogo_poly_w.emf", width = 3, height = 2)
```

# Wire frame black

```{r}
#| eval: false
#| fig-height: 2
#| fig-width: 3

lw = 1.1
col = "black"

glogo_wire_black <- ggplot(dat) +

  geom_line(
   mapping = aes(x=x, y=y),
    colour = col,
    linewidth=lw
  ) +

  geom_line(
    mapping = aes(x=x, y=-y),
    colour = col,
    linewidth=lw
  ) +

  geom_line(
    data = dat |> filter(row_number() <= which.max(y2 > y)),
    mapping = aes(x=x, y=y2),
    colour = col,
    linewidth=lw
  ) +

  geom_line(
    data = dat |> filter(row_number() >= which.max(y3 < y2)),
    mapping = aes(x=x, y=y3),
    colour = col,
    linewidth=lw
  ) +

  geom_line(
    data = dat |> filter(row_number() >= last(which(-y3 < -y))),
    mapping = aes(x=x, y=-y3),
    colour = col,
    linewidth=lw
  ) +

  geom_line(
    data = dat |> filter(row_number() <= last(which(-y2 > -y3))),
    mapping = aes(x=x, y=-y2),
    colour = col,
    linewidth=lw
  )

glogo_wire_black
```

```{r}
#| eval: false
ggsave("gglogo_wire_black.pdf", width = 3, height = 2)
ggsave("gglogo_wire_black.svg", width = 3, height = 2)
ggsave("gglogo_wire_black.emf", width = 3, height = 2)
```

# Wire frame white

```{r}
#| eval: false
#| fig-height: 2
#| fig-width: 3

lw = 1.1
col = "white"

glogo_wire_white <- ggplot(dat) +

  geom_line(
   mapping = aes(x=x, y=y),
    colour = col,
    linewidth=lw
  ) +

  geom_line(
    mapping = aes(x=x, y=-y),
    colour = col,
    linewidth=lw
  ) +

  geom_line(
    data = dat |> filter(row_number() <= which.max(y2 > y)),
    mapping = aes(x=x, y=y2),
    colour = col,
    linewidth=lw
  ) +

  geom_line(
    data = dat |> filter(row_number() >= which.max(y3 < y2)),
    mapping = aes(x=x, y=y3),
    colour = col,
    linewidth=lw
  ) +

  geom_line(
    data = dat |> filter(row_number() >= last(which(-y3 < -y))),
    mapping = aes(x=x, y=-y3),
    colour = col,
    linewidth=lw
  ) +

  geom_line(
    data = dat |> filter(row_number() <= last(which(-y2 > -y3))),
    mapping = aes(x=x, y=-y2),
    colour = col,
    linewidth=lw
  )

glogo_wire_white
```

```{r}
#| eval: false
ggsave("gglogo_wire_white.pdf", width = 3, height = 2)
ggsave("gglogo_wire_white.svg", width = 3, height = 2)
ggsave("gglogo_wire_white.emf", width = 3, height = 2)
```

# Only tops

```{r}
#| fig-height: 4
#| fig-width: 6
glogo2 <- ggplot(dat) + 
  
  geom_line(
    data = dat |> slice(1:(which(y2 > y)[1])),
    mapping = aes(x=x, y=y)
  ) + 
  
  geom_line(
    data = dat |> slice(1:(which(y2 > y)[1])),
    mapping = aes(x=x, y=y2)
  ) + 
  # geom_line(
  #   data = dat |> slice((which(y3 > y2) |> max()):nrow(dat)),
  #   mapping = aes(x=x, y=y3)
  # ) + 
  geom_line(
    data = dat |> slice((which(-y3 < -y) |> max()):nrow(dat)),
    mapping = aes(x=x, y=-y)
  ) + 
  geom_line(
    data = dat |> slice((which(-y3 < -y) |> max()):nrow(dat)),
    mapping = aes(x=x, y=-y3)
  ) + 
  # geom_line(
  #   data = dat |> slice(1:(which(y3 < y2))),
  #   mapping = aes(x=x, y=-y2)
  # )
  NULL

glogo2
```

```{r}
ggsave("gglogo2.svg", width = 3, height = 2)
```

```{r}


dat2 <- tibble(
  x = seq(-3.2, 3.2, length = nx),
  y = dnorm(x),
  y2 = dnorm(x, .58, .75)/2,
  y3 = dnorm(x, .55, .7)/2,
 
  )

dat2
```

```{r}
#| eval: false
#| fig-height: 4
#| fig-width: 6
glogo3 <- ggplot(dat2) + 
  
  geom_line(
    data = dat2 |> filter(row_number() <= which.max(y2 > y)),
    mapping = aes(x=x, y=y)
  ) + 
  
  geom_line(
    data = dat2 |> slice(1:(which(y2 > y)[1])),
    mapping = aes(x=x, y=y2)
  ) + 

  geom_line(
    # data = dat ,
    mapping = aes(x=x, y=y4)
  ) +
# 
#   geom_line(
#     # data = dat |> slice((which(-y3 < -y) |> max()):nrow(dat)),
#     mapping = aes(x=x, y=-y)
#   ) +
# 
#     geom_line(
#     # data = dat |> slice((which(-y3 < -y) |> max()):nrow(dat)),
#     mapping = aes(x=x, y=-y3)
#   ) +
#   
  NULL

glogo3
```

# Discards

```{r}
#| fig-height: 4
#| fig-width: 6
glogo_poly2 <- ggplot() +
  
  geom_polygon(
    data = dat_poly, 
    mapping = aes(x=x, y=y),
    fill = "#00A2E8",
    colour = NA
  ) +
  
  geom_polygon(
    data = dat_poly, 
    mapping = aes(x=-x, y=-y),
    fill = "#00A2E8",
    colour = NA
  ) 

glogo_poly2
```

```{r}
ggsave("gglogo_poly2.svg", width = 3, height = 2)
```



```{r}
#| fig-height: 4
#| fig-width: 6
glogo_poly3 <- ggplot()+
  
  geom_polygon(
    data = dat, 
    mapping = aes(x=x, y=y-min(y)),
    fill = col2,
    colour = NA
  ) +
  
  geom_polygon(
    data = dat, 
    mapping = aes(x=x, y=y2-min(y2)),
    fill = "white",
    alpha = .7,
    colour = NA
  ) +
  
  geom_polygon(
    data = dat, 
    mapping = aes(x=-x, y=-y-max(-y)),
    fill = col2,
    colour = NA
  ) +
  
  geom_polygon(
    data = dat, 
    mapping = aes(x=-x, y=-y2-max(-y2)),
    fill = "white",
    alpha = .7,
    colour = NA
  ) 

glogo_poly3
```

